An integrated distance sampling approach for modelling age-structured populations

### 2023-12-04 Manuscript_Nilsen_Nater.qmd
Edits to manuscript

### 3.2 Case study on willow ptarmigans in Lierne

@@

Having evaluated the overall performance of our model on simulated data, we used data from our case study in Lierne as a case study to estimate abundance, vital rates and detection probabilities from a real-world data set. As is evident from Figure X, the population density increased markedly across the study period, from \< 10 ptarmigan / $km^{2}$ in 2007 to \> 35 ptarmigan / $km^{2}$ in in 2021. The increase was most distinct from 2016 and onward. Across years, mean survival probability for August - January ($S_1$) was estimated at `r filter(tidy_summary_stats, model_parameter =="Mu_S1")$Mean` (95% C.I = `r filter(tidy_summary_stats, model_parameter =="Mu_S1")$lower` - `r filter(tidy_summary_stats, model_parameter =="Mu_S1")$upper`), survival probability for February - July ($S_2$) was estimated as `r filter(tidy_summary_stats, model_parameter =="Mu_S2")$Mean` (95% C.I = `r filter(tidy_summary_stats, model_parameter =="Mu_S2")$lower` - `r filter(tidy_summary_stats, model_parameter =="Mu_S2")$upper`), and annual survival probability $S$ is given by the product of $S_1$ and $S_2$ was estimated at `r filter(tidy_summary_stats, model_parameter =="Mu_S")$Mean` (95% C.I = `r filter(tidy_summary_stats, model_parameter =="Mu_S")$lower` - `r filter(tidy_summary_stats, model_parameter =="Mu_S")$upper`). Note that we assumed that $S$ was constant across all years in the study period from 2007 - 2021. Recruitment ($R_t$) was allowed to vary across years (see model specification), and did as expected show large inter-annual variability. While the mean (baseline) recruitment $\mu_R$ was estimated as `r filter(tidy_summary_stats, model_parameter =="Mu_R")$Mean` (95% C.I = `r filter(tidy_summary_stats, model_parameter =="Mu_R")$lower` - `r filter(tidy_summary_stats, model_parameter =="Mu_R")$upper`).

@@

#### Vital rates

>>>>

### 3.2 Case study on willow ptarmigans in Central Norway

@@

Having evaluated the overall performance of our model on simulated data, we used data from our case study in Lierne as a case study to estimate abundance, vital rates and detection probabilities from a real-world data set. As is evident from Figure X, the population density increased markedly across the study period, from \< 10 ptarmigan / $km^{2}$ in 2007 to \> 35 ptarmigan / $km^{2}$ in in 2021. The increase was most distinct from 2016 and onward. Across years, mean survival probability for August - January ($S_1$) was estimated at `r filter(tidy_summary_stats, model_parameter =="Mu_S1")$Mean` (95% C.I = `r filter(tidy_summary_stats, model_parameter =="Mu_S1")$lower` - `r filter(tidy_summary_stats, model_parameter =="Mu_S1")$upper`), survival probability for February - July ($S_2$) was estimated as `r filter(tidy_summary_stats, model_parameter =="Mu_S2")$Mean` (95% C.I = `r filter(tidy_summary_stats, model_parameter =="Mu_S2")$lower` - `r filter(tidy_summary_stats, model_parameter =="Mu_S2")$upper`), and annual survival probability $S$ is given by the product of $S_1$ and $S_2$ was estimated at `r filter(tidy_summary_stats, model_parameter =="Mu_S")$Mean` (95% C.I = `r filter(tidy_summary_stats, model_parameter =="Mu_S")$lower` - `r filter(tidy_summary_stats, model_parameter =="Mu_S")$upper`). Note that we assumed that $S$ was constant across all years in the study period from 2007 - 2021.

Recruitment ($R_t$) was allowed to vary across years (see model specification), and did as expected show large inter-annual variability. While the mean (baseline) recruitment $\mu_R$ was estimated as `r filter(tidy_summary_stats, model_parameter =="Mu_R")$Mean` (95% C.I = `r filter(tidy_summary_stats, model_parameter =="Mu_R")$lower` - `r filter(tidy_summary_stats, model_parameter =="Mu_R")$upper`). The estimated recruitment rate varied (as expected) among years: while the annual recruitment rate was estimated at `r max(filter(tidy_summary_stats, model_parameter =="R_year")$Mean)` in year 2007, it was estimated at `r min(filter(tidy_summary_stats, model_parameter =="R_year")$Mean)` in year 2012. Moreover, as expected based on the Alternative Prey Hypothesis (APH), our model estimated a strong link between the annual recruitment and the abundance of small rodents (slope-paramater for the z-standardized rodent occurrence data = `r filter(tidy_summary_stats, model_parameter =="betaR.R")$Mean` ; 95% C.I. = `r filter(tidy_summary_stats, model_parameter =="betaR.R")$lower` - `r filter(tidy_summary_stats, model_parameter =="betaR.R")$upper`).

@@
